---
title: About
author: Ben Harden
date: '2021-05-20'
slug: []
categories: []
tags: []
draft: no
---

The PEP Class is built to help prepare you for the research work you will be doing this summer. This includes allowing you to:

- Build your knowledge of broad oceanographic and environmental science topics
- Develop you research skills
- Understand the community you are becoming a part of

The class will begin on June 1 and run until June 25.

## Class Philosophy

Before I get to details, I want to start by laying out my philosophy for this class. 

You are all coming to this class with different experiences. Some of you have had previous research opportunities, some none. Others will have been exposed to coding skills, many will not. Some of you will be environmental science majors, while for others this will be a new topic to grapple with. Regardless of your previous experiences, you all have strengths to bring to the table. You all have a responsibility to keep your eyes open for your peers who are struggling and help them out. This class is a **collaborative space** so that everyone can grow their skills and understanding as much as possible. No one succeeds because someone else fails. We all grow together.

The class will be mostly **self-paced**, meaning you are free to move through the content and exercises when it suits you. My reasons for this are to help you:

* Plan your time to best suit your learning style and energy levels
* Do work in timezone when others (including me) will not be free/working
* Have more time for things you struggle with by allowing you to go faster through things you might find more straight-forward
* Have more agency about how you spend your time and take breaks when you need them
* Not get too focused keeping track of many deadlines
* Collaborate with your peers at times that suit you
* Prepare for your research which will be intrinsically self-paced

The only scheduled parts of the class will be various daily discussions, group meetings, check-ins, and guest speakers. More on the structure below.

I value feedback and self-reflection. You will be given many opportunities to let me know how you are getting on, what you are understanding, and what you have questions about. My hope is that through this process that you become conscious of your strengths and areas for growth. You will also be helping me understand how best to help you and your fellow students. For the duration of the PEP class, you are my primary professional activity so feel empowered to contact me or the Course Assistant if you have any questions or concerns.

Finally, remember throughout: you belong here at PEP and in the Woods Hole science community. The PEP staff are here to support you and help you be successful.

## What we expect from you

We hope you will see this class as more than simply a requirement to gain a grade for college credit. It should be a useful tool for you to advance your thinking and skills around conducting research in Environmental Sciences. It should be directly relevant to the work you will be doing this summer as well as building a broader tool set for many careers in science.

We know that this has been a long year of digital learning for many of you. We are aware of the limitations and know that it is not ideal for so many reasons. We have tried to develop this class in a way that will limit the one-way delivery of information over video call. However, we still need you to find the motivation to once more enter a digital world of learning. We will be bringing our energy and we expect you to too. 

We expect you to use your time in the PEP class effectively. The self-paced mode will give you lots of flexibility to work when and how you like, but we do expect you to engage with all aspects. Work should take between 3 and 4 hours a day and work at the weekend is likely to complete certain assignments. Use your peers as a support network for things you are finding hard and to keep you accountable. 

We welcome your feedback (and will ask for it frequently) so please let us know how you are getting on and what isn't sticking and we will try and help you out. Please also let us know how long the classwork is taking you so we can make sure

## What you can expect from your instructors

You instructors will be available for immediate feedback and discussion during the daily class period (12-3pm, see below). Please speak up if you need our support, but be patient as we try to meet the needs of the whole PEP cohort. If there are enough questions or interest in a particular topic then we will arrange small group instruction.

Outside the class window we will also be available to answer questions and give feedback. Please contact us through Slack and we will get back to you as soon as we are able. Again, please be patient.

## Class Structure

OK, so on to the nuts-and-bolts. 

We're going to be looking at four topic areas:

* **Oceanography:** The study of the world's oceans from physical, chemical, biological, and geological viewpoints.
* **Research Skills:** Finding and Reading literature, exploring data, collaboration, poster/oral presentations.
* **Data Visualization:** Technically a subtopic of research skills, but one with enough weight to give it's own topic.
* **Advanced Topics**: Invited guests from the Woods Hole science community to expand your understanding on important topics and introduce you to the institutions in Woods Hole.

Each topic will be broken down into a series of modules, each dealing with a portion of the wider topic. In each module there are a series of levels to complete. You can see the full list of topics, modules and levels on the [classes](../classes/) page.

### Levels and Grades

There will typically be four levels in any module. These need to be completed in order with an assignment/assessment at the end of each level that needs to be completed before you can move on. The level pages themselves will provide clear instructions for what you have to do, but as an overview:

```{r, echo = F}
knitr::kable(
  tibble::tribble(~"Level", ~"Typical Work", ~"Assessment", ~"To move on...", ~"Module Grade",
               "1", "Reading some basic information","Multichoice quiz (scored 0-5)", "Score 4 or more", "75%",
               "2", "Appling ideas to new situation", "Written/spoken assignment", "Review and respond to feedback", "84%",
               "3", "Appling and Analysing a problem", "Various assignments (scored 0-8)", "Must score 5 or more","84% + level 3 score",
               "4", "Evaluating or creating", "Various assignments (scored 0-8) ", "N/A","84% + level 3 score + level 4 score")
)
```

I am expecting everyone to engage with every module to Level 3. Level 4 is advanced work and is optional. You do not need to complete all Level 4 work - you are free to skip it on any module (but you will probably need to complete some if you want to maintain an A average). You are free to get feedback and revise any assignment up until the end of the class - the gradebook will stay open until July 2.

*NOTE: The Advanced Topic will not have level 4. Grades will be Level 1 (pass/not-pass 80%), Level 2 (pass/not-pass 90%), Level 3 (optional, scored 0-10 -> 100%)*

When you have completed the work from a level, you need to let us know by clicking the "Level Complete" button at the bottom of the level page and giving a small amount of reflection/feedback. This will notify us that there is work that needs to be reviewed. Until you get feedback, you are free to move on to another topic/level.

Your total class grade will be calculated with the following weights:

```{r, echo = F}
knitr::kable(
  tibble::tribble(~"Topic", ~"Weight",
                  "Oceanography", "20%",
                  "Research Skills", "40%",
                  "Data Visualization", "20%",
                  "Advanced Topics", "20%")
)
```

### Staying On-pace

You are free to move at a pace that suits you for the most part. However, there will be a number of fixed-point assessments, discussions and presentations. These will be generally tied to the completion of level 3 work and may be collaborative - you should keep an eye on the dates/times of these so you can make sure you move through your levels in time to fully participate.

To help you there will be a **public tracker** that will show which levels you and your peers have completed. It will also have a "pace-setter" - my sense of where you should be in any module on any day to be ready for the fixed-point activities. Please use this to help you gauge what you might need to do next and to see which of your peers might be able to help you with work if you are stuck.

### Class Interaction

Class will be 12pm to 3pm Eastern Daily. This is a broad period in which we'll have both synchronous and asynchronous class work. You instructors will be available to you during this time and we expect you to be available for full-class, small group and individual interaction as needed.

Class will begin every day at 12pm with a short survey and check-in session. This space will be used for announcements, making a plan for the rest of class, and gathering topics that might need small-group or individual attention.

After the initial full-group check-in will be free time to work. We will be available at this time for questions and assistance and will also be scheduling in small group check-ins, one-on-ones and other activities at this time to support students with their work.

During free-work periods, you should feel empowered to take short breaks, grab some food, step outside to get some air, etc. However, this is our time to work with you and we expect you to be available to engage with us if we need to get hold of you.

2pm to 3pm will be designated as the time for full-group assessments, discussions and presentations. Please be fully available during this time if there is something scheduled. The calendar of these events will be kept current in the PEP 2021 Google Calendar.

Outside of the 12pm-3pm window, your instructors will also be available to field questions and offer assistance from 9am to 5pm Daily. Please post your questions in the #class Slack Channel and be patient for a reply. Remember, your peers will often be a great resource for you - please work together if you like, but remember to only submit your own work (see policy on academic accuracy in the syllabus).

## Platforms

To deliver the class content and skills, we will be using the following platforms. I have indicated below which of these you will have to sign-up for.

### Class Site

The site you are reading this on is the class website. It will be where we share the work you will need to complete for each level, module and topic. If a submission is required I will link directly to it in Google Classroom or RStudio Cloud (see below).

### Google Classroom

Google Classroom will be where we organize some assignments and submissions for various levels. You will need to sign-up for the class with the link and access code I will send to you through Slack.

### Slack

To facilitate direct communication with and between members of the 2020 PEP Cohort, we will be using Slack. This is a direct and group messaging application that will allow us to share news and upcoming events with the whole group as well as organize smaller discussions or contact individual students. You should already have signed-up for a Slack account.

Please contact us through Slack if you have questions or need assistance.

### Zoom

We will be using Zoom for the synchronous part of classes. I will schedule the zoom meetings and post them on Slack and on the PEP calendar. There is also a link to the meetings on Google Classroom.

### Wonder

We are experimenting this year with using [Wonder](https://www.wonder.me/) to organize more fluid discussion and class time. This is a video and DM web app which will allow us to gather small groups as needed to talk over specific topics. There will be a wonder classroom set up for the duration of the class and we expect you to be there during the class free time so we can get a hold of you. You can also use this space to work with others outside of class time. The link to this classroom will be in Slack and on Google Classroom.

### RStudio Cloud

We are going to be using a programming language called R for our data visualization work. To make this easy for me to help you, and for you to share your code with me, we are going to be doing this work on a cloud version of the software. You will need to go to <https://rstudio.cloud/> to sign up for a free account. I will then send you a link to my classroom for you to join. We'll talk more about how to use this platform in class.

### Google Drive

You will have also received an email about a Google Drive folder that has been set up for you at SEA. Please use this to store any work related items that you need in one place or want to share with others. This is also a good place to back-up your work. We will also be putting various documents in here for you if we need to.

## Resources/Texts

Our general oceanography text will be:

[Introduction to Ocean Sciences](https://www.reefimages.com/oceansci.php), Second digital edition, ver 4.14 by Douglas A. Segar

This is a good introductory text on the many aspects of ocean science and is all [available online](https://www.reefimages.com/oceansci.php) for free. We'll be directing you to readings by chapter number and page. In addition, there is a section called [Critical Concepts](https://www.reefimages.com/oceans/SegarOcean4CritCon.pdf) which covers ideas that span multiple chapters. We will be directing you to review some of those by number also. The chapters all load as PDFs so you might consider saving assigned chapters to your computer and organizing them for future browsing.

For R visualization work, the core text is:

[R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. This is a great intro text to working with data in R. I recommend reading the whole thing over the course of the summer, but I'll be assigning specific chapters in class.

Other readings and resources for the class will be distributed as required. 




