---
title: About
author: Ben Harden
date: "`r format(Sys.time(), '%Y-%m-%d')`"
slug: []
categories: []
tags: []
draft: no
---

This Data Communication and Visualization class will help prepare you for the data research and communication work you will be doing in your future careers by building skills in: 

- Data Visualization in the R programming language
- Understanding good data and visual design strategies
- Optimizing your verbal communication of data

The class will begin on October 4th, 2021 and run until the end of the program on December 23rd, 2021, but the majority of the classwork will be concluded by the end of the shore component (November 15th, 2021).

## Class Philosophy

Before I get to details, I want to start by laying out my philosophy for this class. 

You are all coming to this class with different experiences. Some of you have had previous research opportunities, some none. Others will have been exposed to coding skills, many will not. Some of you will be oceanography majors, while for others this will be a new topic to grapple with. Regardless of your previous experiences, you all have strengths to bring to the table. You all have a responsibility to keep your eyes open for your peers who are struggling and help them out. This class is a **collaborative space** so that everyone can grow their skills and understanding as much as possible. No one succeeds because someone else fails. We all grow together.

The class will be mostly **self-paced**, meaning you are free to move through the content and exercises when it suits you. My reasons for this are to help you:

* Plan your time to best suit your learning style and energy levels
* Do work in timezone when others (including me) will not be free/working during initial virtual component of the program.
* Have more time for things you struggle with by allowing you to go faster through things you might find more straight-forward
* Have more agency about how you spend your time and take breaks when you need them
* Not get too focused keeping track of many deadlines
* Collaborate with your peers at times that suit you
* Have the most effective in-class experience wherein we can focus on the needs of small-groups and the class as a whole based on your work to date.

Alongside the self-paces work will be scheduled portions of the class that will include group work, discussions, presentation, etc. More on the structure below.

I value feedback and self-reflection. You will be given many opportunities to let me know how you are getting on, what you are understanding, and what you have questions about. My hope is that through this process that you become conscious of your strengths and areas for growth. You will also be helping me understand how best to help you and your fellow students. For the duration of the class, you are my primary professional activity so feel empowered to contact me if you have any questions or concerns.

## What I expect from you

I hope you will see this class as more than simply a requirement to gain a grade for college credit. It should be a useful tool for you to advance your thinking and skills around communicating data effectively. These skills are applicable to so many careers, not just for those in academic sciences.

I expect you to use your time in the class effectively. The self-paced mode will give you lots of flexibility to work when and how you like, but I do expect you to engage with all aspects. We will meet most days of the week for a total of 7 hours of weekly contact time. You should expect to spend a comparable amount of time on your work out of class. Use your peers as a support network for things you are finding hard and to keep you accountable. 

I welcome your feedback (and will ask for it frequently) so please let me know how you are getting on and what isn't sticking and I will try and help you out. Please also let us know how long the classwork is taking you so I can make sure I am calibrating the quantity and depth of content accordingly.

## What you can expect from me

I will be available for immediate feedback and discussion during the near-daily class period (see below) and during out of class office hours. Please speak up if you need specific support during any of these times - the more heads-up I can have the more effective I can be at helping you. However, please be patient as I try to meet the needs of the whole class. If there are enough questions or interest in a particular topic then I will arrange small group instruction.

Outside the class window I will also be available to answer questions and give feedback. Please contact me by email which I check daily from 8am to 8pm except for Saturdays.

## Class Structure

### Topics, Modules and Levels (and Grades)

The class content on this site is organized into three topics:

* Visualize: learning how to effectively use R to create compelling graphics
* Communicate: Increasing your communication capacity through radio storytelling
* Design: Introduction to key elements of data design and data presentations


Each topic is comprised of a set of modules, each with four levels of work. Topics do not need to be completed in order, but modules and levels are typically sequential with an assignment/assessment at the end of each level that needs to be completed before you can move on.

#### Modules

The level pages in a module will provide clear instructions for what you have to do and you should move through the levels of each module sequentially.

I am expecting everyone to engage with every module to Level 3. You do not need to complete all (or any) Level 4 work - you are free to skip it on any module (but you will probably need to complete some if you want to maintain an A average - see below). You are free to get feedback and revise any level assessment up until the end of the class - my grade book (other than your final at-sea assignments) will stay open until you board the ship on November 15th.

When you have completed the work from a level, you need to let me know by clicking the "Exit Ticket" button at the bottom of the level page and giving a small amount of reflection/feedback. This will notify me that there is work that needs to be reviewed. Until you get feedback, you are free to move on to another topic/level.

Grading-wise, the typical grade structure for most modules is as follows:

```{r, echo = F}
knitr::kable(
  tibble::tribble(~"Level",~"To Move on...", ~"Module Grade",
               "1", "Sign-off by instructor", "70%",
               "2", "Sign-off by instructor", "80%",
               "3", "Score 6 or more", "80% + level 3 score (out of 10)",
               "4", "N/A", "80% + level 3 score + level 4 score (out of 10)")
)
```

Levels 1-2 are pass/not-pass whereas levels 3 and 4 are normally scored.

*NOTE: The one exception to this structure is in the Visualize modules (coding). The scores for each level will be weighted differently depending on your previous exposure to this topic. If you have no or limited experience coding, the score allocation will be L1: 80%, L2: 90%, L3: 100%, but if you have previous experience, I'd like to see you push yourself and complete more L4 work to get the same score - I will discuss this with you individually.*

#### Grading

Your final grade will be calculated based on the weighted scores in each module. Each standard module carries the same weight with project modules carrying twice that weight. As a result, your total class grade will be calculated with the following weights:

```{r, echo = F}
knitr::kable(
  tibble::tribble(~"Topic", ~"Module Content", ~"Module Final Project",
                  "Visualize", "22.5%", "15%",
                  "Communicate", "15%", "15%",
                  "Design", "7.5%", "15%",
                  "Total", "45%", "45%")
)
```

The final 10% of your grade will come from a final at-sea assessments of your DCV skills through a contribution to the cruise story map, an example of which can be found [here](https://storymaps.arcgis.com/stories/8ec1591c831e4d568110c0d5f65bcbf7).


### Staying On-pace

You are free to move at a pace that suits you for the most part. However, there will be a number of fixed-point assessments, discussions and presentations. These will be generally tied to the completion of level 3 work and may be collaborative - you should keep an eye on the dates/times of these so you can make sure you move through your levels in time to fully participate.

To help you there will be a **public tracker** that will show which levels you and your peers have completed. It will also have a "pace-setter" - my sense of where you should be in any module on any day to be ready for the fixed-point activities. Please use this to help you gauge what you might need to do next and to see which of your peers might be able to help you with work if you are stuck.

### In-preson interaction

Class will be at the following times:

* Monday: 10:45 - 12:15
* Tuesday: 13:30 - 15:45
* Thursday: 10:45 - 12:15
* Friday: 10:45 - 12:15

During class time there will typically be a work block time as well as time set aside for questions, group activities and discussions. Those that form assessments will be clearly posted and labeled.

Outside of the class periods I will be available to field questions and offer assistance over email or during office hours:

* Tuesday 09:00 - 10:30
* Thursday 09:00 - 10:30
* Everyday 12:15 - 13:30

Remember, your peers will often also be a great resource for you - please work together if you like, but remember to only submit your own work (see policy on academic accuracy in the syllabus).

## Platforms

To deliver the class content and skills, we will be using the following platforms. I have indicated below which of these you will have to sign-up for.

### Class Site

The site you are reading this on is the class website. It will be where I share the work you will need to complete for each level, module and topic. If a submission is required I will link directly to it in Google Classroom or RStudio Cloud (see below).

### Google Classroom

Google Classroom will be where we organize some assignments and submissions for various levels. You will need to sign-up for the class with the link and access code I sent to you by email,

### RStudio Cloud

We are going to be using a programming language called R for our data visualization work. To make this easy for me to help you, and for you to share your code with me, we are going to be doing this work on a cloud version of the software. You will need to go to <https://rstudio.cloud/> to sign up for a free account. I will then send you a link to my classroom for you to join. We'll talk more about how to use this platform in class.

## Resources/Texts

See the syllabus for a full list of recommended texts, most of which are available in the SEA library if I haven't linked you to digital versions.




