---
title: Vizualize SAQs
author: Ben Harden
date: '2021-10-05'
slug: []
categories: []
tags: []
draft: no
---

## Sometimes Asked Questions

### What are git, error trapping, and debugging?

These expressions all came up in [Section 1.3](https://robwschlegel.github.io/Intro_R_Workshop/prelim.html#why-use-r) of the Intro to R page I asked you to read.

* **git**: Is a version tracking software. Most people who work with code (and especially in a team) find using version tracking software useful. It allows you to reclaim code lines you had previously deleted, split off a new version of your code for trialing before integrating it back in, and allows you to sync work between computers. It's basic use works a lot like what you find on Google Docs where you can see the edit history and return to a previous state as needed.
* **Error Trapping**: This is the process by which you can build in elements of your code that will test to see whether the outcome has "worked" or produces something useful for you code to continue. For example, you might get into a situation where the result of some calculation doesn't produce a number (it creates an "NA" or "Not Available" entry) that the rest of your code is relying on. Rather than let the code play out and make a mess, you can "trap" this error and either modify how your code continues or stop your code and send an error message to the console.
* **Debugging**: The process by which your code wont work the way you want it to and you will have to go back through the code to find the "bug" or incorrect use of a function and correct it.

### What is the difference between a double and an integer? Why would we use integers?

A double is a decimal number (i.e. 4.56) and an integer is a rounded whole (1, 2 ,3, etc.). R is pretty flexible about which is which compared to other languages. The main difference for other languages where speed and space might be an issue is that storing integers takes less space and are can be recalled quicker. This wont be a problem for the kind of work we're doing, but it's at least good to be aware of.

### Are there any usability differences/drawbacks in the cloud vs. normal RStudio?

You have the same functionality on the cloud as you do with desktop RStudio. The only difference might be in performance - the cloud has to send your commands to a computer somewhere else, run the code, and then send the output back to you. It may run just a little slower as a result. The up-side for this class though is that I have access to all your work and can jump in to check how you're doing, add comments/suggestions, and help debug your code as needed. You can also share your projects with your peers so they can see your work too.

### Can you pick some columns not to load with the csv, or do you have to remove them after the dataset is loaded?

You *can* choose specific columns to load in. See the help file for `read_csv()` by entering `?read_csv` into the console and pressing enter. These help files exist for all functions in R and explain how each function works and what arguments (choices) are available to you. In this case, the `col_types` argument will allow you to select only certain columns, but unless you data set is HUGE then in my experience it's almost always easier just to read in all of it and then slim down after.

### Is R a combination of programs such as python and excel or does it lean more towards one of these in terms of function?

R is it's own thing - a language unto itself. It has similarities to both python (another widely used, accessible programming language) and Excel, but operates independently of both. In terms of how it functions, the coding nature of R leans it towards python and away from the point-and-click, visual processes in Excel. So, how do R and Python compare? Python is more flexible than R - you can use it more easily for all manner of tasks on a computer and set it to task organizing files, running monitoring devices, etc. (I use it for running my home weather station), but R is still the go-to for data analysis and visualization, which is built into python as an after-thought and doesn't have the same level of completeness or community support. R was built specifically for working with data.

### What is the difference between data types integer and character?

Integers are whole numbers (e.g. 1, 5, 75, etc.) and characters are words (e.g. "hello", "testing", "this thing with spaces", etc.)

### Is the syntax in R similar to any other language, which is it closest to?

Each language is unique in some sense. R probably skews closest to Python in how it looks (but also MATLAB and Java) - vectors are indexed using [ ], functions are run with ( ) containing the arguments. However, there are also some big differences that I am happy to discuss with anyone interested. Another similarity between R and Python are that they are both interactive languages - you can write scripts but also enter commands directly in the console. This differs from other languages (such as C++ and FORTRAN) where the only way to implement code is through scripting.

### Does the data for the data frames have to be entered in R or can it be imported from external sources or programs?

You can indeed enter data manually into R, but the most efficient process is to read in the data from some external source. We have seen one example in [Module 1, Level 3](../../classes/visualize/01-introduction/level-3/) of how to import data from a Comma Separated Value file (pretty common data format), but there are also functions in R to read pretty much any other data format you care to think of - NetCDF, JSON, TSV, Shapefiles, XML, YAML, HTML, JPEG, GeoTIFF, etc.

### What is the difference between commands like count, sum, tally, and n? 

You'll often come across different functions in R that might end up doing roughly the same thing, especially once you start adding in different packages. You should always feel empowered to find out more about how a function works by writing `?<function-name>` into the console. A help file will show up to the right which will tell you all about the function and give you some examples of how to use it (usually the most helpful part to start with). Each function has a list of "arguments" which are the things you need to (or can optionally) put into the function to get the output. Similar functions like the above will probably have different arguments and may rely on a different process under the hood (or bonnet as I might say).

### How do I create a csv file?  Do I do it within R or do we import the data as a csv file from excel or another similar software?

OK, so having dismissed Excel as a robust data science tool, I have to say that for entering data, Excel is still one of the best tools out there. To create a csv file, I often enter the data as a table in Excel and then save the file as a CSV file and then read it into R to analyze. That said, it's not everyday that you'll need to do this as that method only really applies to hand-entered, new data. Most data we'll use in this class has come from ocean instrumentation and satellite data and you'll download it as a CSV file directly form the web. This will remove Excel from the picture entirely.

### What is the difference between read.csv() and read_csv()?

So many of you will probably have come across this slight snag. RStudio is super helpful a lot of the time - it "suggests" autocompletes for functions and objects once you begin typing. You can also hit the tab key when you start writing a command and it will give you options for how you might want to complete what you are writing. That said, this feature can often lead to you selecting the wrong (but oh so similarly sounding) function. 

In this case, `read.csv()` is the baseR or "original" csv reading function. It works much the same as the `read_csv()` function that comes from the `readr` package, but it is typically slower and produces some odd things in the output that most people typically don't want it to do. You should get into the habit of using `read_csv()` to avoid potential issues down the line. For more info, there's a good description of the differences [here](https://r4ds.had.co.nz/data-import.html#compared-to-base-r).

### What is the difference between a vector and a list?

A vector is a series of numbers or words contained in a single object. You will most likely see the following as ways to make vectors in R:

* Using the c() function to "combine" entries

```{r}
c(1,2,5,75,43,234.3)
```

or

```{r}
c("one potato", "two potato", "three potato", "four")
```

* Using the `:` feature to make a vector of sequential numbers

```{r}
3:20
```

* Using the seq() function to make a sequence of non-sequential numbers

```{r}
seq(1, 10, by = 0.2)
```

Lists on the other hand look similar to vectors in some regards, but you can put something of any size in each component of a list:

```{r}
list(1,2:3,"hello",tibble::tibble(x = 1:4, y=3:6))
```

Don't worry too much about lists - they are a really useful tool if you want to put things of different types together, but we wont be coming across them in this class. 


### How can you tell what the units are for data that was inputted into a table in R? 

This will depend on the data set. Typically, best preactice is that units are not encoded into the data itself, but supplied in a separate metadata file. Thats basically what I'll be providing you with - the data, but also a description elsewhere of what the data is in each column and the units. This *can* get a little messy though if the metadata and data files are separated. that's why in some cases you'll see the units in the column names, or as the first row in the data, or supplied in some header that sits above the data. I prefer keeping the metadata separate (as many people do) becuase it makes reading the data in so much more simple if you don't have to remove headers and edit our units from column names.

### What is the difference between filter/select and using $ ?

Filter is when you just want to select certain rows from your data which meet a certain criteria. Maybe they all have the same value in one column. Maybe you want all the rows where the wind speed is greater than 10 m/s. Whatever the criterion, you can use filter to acheive this goal of isolating rows.

```{r, echo = F, message=FALSE, error=F, warning=FALSE}
library(tidyverse)
```


```{r}
filter(mtcars, cyl == 6)
```


Select just takes certain columns from your data. This is similar to using the `$` operator, but allows you to select more than one row.

```{r}
select(mtcars, cyl, disp)
```

### At the end of the data, the last lines tells you how many more rows and variables are not shown, then how can we see the rest of the data? Can we only see it when we use the table icon on the environmental tab?

You *can* have R print out more rows, but it's often best to use:

```{r, eval = F}
View(mtcars)
```

or click the table icon to get a full view of the data.

You can also use:

```{r}
print(mtcars)
```
to see all the data in the console.
