---
title: 'Annotations and Themes'
author: Ben Harden
date: '2021-05-05'
level: 2
slug: []
categories: []
tags: []
draft: no
quiz: https://rstudio.cloud/spaces/170089/project/2870358
---

```{r, message=F, warning=F, echo=FALSE}
library(tidyverse)
```


In this level, we've going to cover two parts of the ggplot package that are more geared to data communication and design - annotations and themes. Annotations (or labels) are the text that lives on your plot that helps someone (or a future you) interpret what is displayed. A theme is ggplot's term for the data design elements of the figure. This covers all the things you could think about changing for your non-data elements of the plot, e.g. grid lines, label orientation, background color, axis lines, axis ticks, font size/type, etc. 

## Labels and Themes Primer

You should work through the "Labels" and "Themes" section of [this primer](https://rstudio.cloud/learn/primers/3.8) to get the basics of each under your belt.

### Additional notes on Labels

The labels section the Primer above really only talks about making a title, subtitle, and caption for your plot. In many situations, you probably wont actually be setting any of these - figures in papers, for example, have the caption as separate text and a title isn't included as part of the figure. That said, the `labs()` function is still very useful in setting labels for other parts of the plot.

Let's take the 5 CTD profiles we plotted in our exercise back in our [intro to plotting module](../../02-plotting1/level-2).

```{r}
# Reading in the profile data from the 5 CTD casts
profiles <- read_csv("profiles.csv")

# plotting all temperatures verses depth
# Assigning this plot to a variable for use in following sections
p <- ggplot(profiles, aes(temp, z)) +
  geom_path(aes(color = Cruise)) +
  scale_y_reverse()

# Making the plot appear
p
```
We can now use labs to change our x, y and color axes labels in the same way we'd set the title, subtitle, etc.

```{r}
# plotting all temperatures verses depth and labeling axes
# I'm overwriting p here to add labels to the object
p <- p +
  labs(x = "Depth (m)",
       y = "Temperature (°C)",
       color = "Cruise ID")

# Making the plot appear
p
```

*NOTE: In this code chunk, I've inserted a return after each entry in the labs function - R can interpret these line breaks as a continuation of the function as long as they come after a comma. This can often help your code look more organized*

You can also get rid of a label by setting it's value to `NULL`

```{r}
p <- p +
  labs(x = "Depth (m)",
       y = "Temperature (°C)",
       color = NULL)
```

## Exercise

Head over to the RStudio Cloud assignment test your understanding of how to edit these features of a plot.